---
title: "Assignment B-1: Making a function"
author: "Colton Stevens"
output: github_document
---
Load the relevant packages.
```{r, message = FALSE}
library(tidyverse)
library(gapminder)
library(testthat)
```

## Exercise 1: Make a Function (25 points) and Exercise 2: Document your Function (20 points)

In this exercise, you’ll be making a function and fortifying it. The function need not be complicated. The function need not be “serious”, but shouldn’t be nonsense.

In the same code chunk where you made your function, document the function using roxygen2 tags. Be sure to include:

  1. Title.
  2. Function description: In 1-2 brief sentences, describe what the function does.
  3. Document each argument with the @param tag, making sure to justify why you named the parameter as you did. (Justification for naming is not often needed, but we want to hear your      reasoning.)
  4. What the function returns, using the @return tag.

This function takes in a dataframe and a numerical column in that dataframe and outputs a data.frame of summary statistics (mean, median, max, min, standard deviation, and count), as well as a histogram of that column's distribution.
```{r}
#' Summary Statistics for a Numeric Column in a Dataframe
#' 
#' This function takes in a dataframe and a numerical column in that dataframe and outputs a data.frame of summary statistics (mean, median, max, min, standard deviation, and count), as well as a histogram of that column's distribution with a vertical line representing the mean.
#' @params data The data.frame to be inputted. This is named `data` because it describes what the first argument should be — a data.frame.
#' @params col The numerical column to be used for calculating summary statistics. This is named `col` because it describes what the second argument should be — a column.
#' @return A data.frame of summary statistics (mean, median, max, min, standard deviation, and count), as well as a histogram.
#' @examples
#' summary_stats_column(gapminder,lifeExp)
#' summary_stats_column(mtcars, mpg)

summary_stats_column <- function(data, col) {
  calculations <- dplyr::summarise(data,
                                   is_numeric = is.numeric({{col}}),
                                   class = class({{col}}))
  if(!calculations$is_numeric){
    stop("Selected column is not numeric. Column is: ", calculations$class)
  }

  summary_table <- data %>%
    dplyr::summarise(mean = mean({{col}}, na.rm = TRUE),
            median = median({{col}}, na.rm = TRUE),
            max = max({{col}}, na.rm = TRUE),
            min = min({{col}}, na.rm = TRUE),
            st_dev = sd({{col}}, na.rm = TRUE),
            count = n())
  
  summary_plot <- ggplot(data) +
    geom_histogram(aes({{col}}),
                   bins = 30) +
    geom_vline(aes(xintercept = mean({{col}})), 
             linetype = "dashed", size = 0.6)
  
  output <- list(summary_table, summary_plot)
  return(output)
}
```
## Exercise 3: Include examples (15 points)

Demonstrate the usage of your function with a few examples. Use one or more new code chunks, describing what you’re doing.

Here are some examples. The first example `summary_stats_column(gapminder, lifeExp)` creates a summary statistics data.frame and histogram of the `lifeExp` column in the `gapminder` data.frame.

```{r}
summary_stats_column(gapminder, lifeExp)
```
The second example `summary_stats_column(mtcars, mpg)` creates a summary statistics data.frame and histogram of the `mpg` column in the `mtcars` data.frame.

```{r}
summary_stats_column(mtcars, mpg)
```

The third example `summary_stats_column(gapminder, country)` attempts to create a summary statistics data.frame and histogram of the `country` column in the `gapminder` data.frame, but it throws an error and stops because the `country` column is not numeric.

```{r error = TRUE}
summary_stats_column(gapminder, country)
```


## Exercise 4: Test the Function (25 points)

Running examples is a good way of checking by-eye whether your function is working as expected. But, having a formal “yes or no” check is useful when you move on to other parts of your analysis.

Write formal tests for your function. You should use at least three non-redundant uses of an expect_() function from the testthat package, and they should be contained in a test_that() function (or more than one). They should all pass.

```{r}
test_that("The output is a list; the first item of the output is a data.frame, the second item of the output is a ggplot, and neither item is empty", {
  expect_true(is.list(summary_stats_column(gapminder, lifeExp)))
  expect_true(!is.null(summary_stats_column(gapminder, lifeExp)[[1]]))
  expect_true(!is.null(summary_stats_column(gapminder, lifeExp)[[2]]))
  expect_s3_class(summary_stats_column(gapminder, lifeExp)[[1]], "data.frame")
  expect_s3_class(summary_stats_column(gapminder, lifeExp)[[2]], "ggplot")
  })

test_that("Error occurs for incorrect column arguments (a non-numeric column and a NA column)", {
          expect_error(summary_stats_column(gapminder, country))
          expect_error(summary_stats_column(gapminder, NA))
  })

test_that("The first input is a dataframe", {
  expect_true(is.data.frame(gapminder))
  })
```

